---
title: "Inferential Analysis of effect of Vit. C on Tooth Growth"
author: "Odin Matanguihan"
date: "May 2, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

In this report, toothgrowth data is analyzed to draw inferences.  t.test() is used to analyze correlation between vit. C and tooth growth.  It is analyzed by intake type and dosage.



```{r ToothGrowth}

data(ToothGrowth)   #load the data
head(ToothGrowth)
str(ToothGrowth)
```

#Provide basic summary of data

```{r}

ToothGrowth$dose <- as.factor(ToothGrowth$dose)
table(ToothGrowth$supp, ToothGrowth$dose)

```


A simple boxplot already gives a good idea of what the conclusions are going to be.

```{r pressure, echo=FALSE}
library(ggplot2)
plot <- ggplot(ToothGrowth, aes(x=factor(dose),y=len,fill=factor(dose)))
plot <- plot + geom_boxplot()  + facet_grid(.~supp)
plot <- plot +  scale_x_discrete ("Dosage(in mg)") + scale_y_continuous("Tooth Length") 
plot <- plot +  ggtitle("Effect of Dosage and Supplement Type on Tooth Growth")
plot
```

The plot above shows that tooth growth is affected by dosage, and that OJ(orange juice) is better than VC(ascorbic acid).  The difference seems to disappear at the higher dosage of 2 mg.  We can confirm if there is a difference using a t test.  The null hypothesis here is that there is no difference.  

```{r}

x <- t.test(len ~ supp, paired=FALSE, var.equal=FALSE, data=ToothGrowth[ToothGrowth$dose == 0.5, ])
x$conf.int
x$estimate

```

The confidence interval does not include zero and the null hypothesis is rejected for dosage = 0.5 mg.

```{r}

x <- t.test(len ~ supp, paired=FALSE, var.equal=FALSE, data=ToothGrowth[ToothGrowth$dose == 1, ])
x$conf.int
x$estimate

```

For the dosage = 1 mg, the null hypothesis is also rejected.

```{r}

x <- t.test(len ~ supp, paired=FALSE, var.equal=FALSE, data=ToothGrowth[ToothGrowth$dose == 2, ])
x$conf.int
x$estimate

```

At dosage = 2 mg, the 95% confidence interval includes zero right in the middle.  The difference in means between the two vit C delivery sources is not statistically significant.

# Conclusions

1.  Higher dosages of vit. C leads to improved tooth growth.
2.  OJ gives better tooth growth than VC.
3.  The difference in tooth growth becomes insignificant at 2 mg. dosage

# Assumptions

1.  The variances between sample populations are not equal.
2.  The sample data is not paired.